Weight sharing via comonads

### 2020-10-08 backup/presimpler/main.tex
before volterra

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Fun}{{\textnormal{Fun}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Set}{{\mathbf{Set}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\DCat}{{\mathbf{D}}}
\newcommand{\FCoalg}{{F\textnormal{-coalg}}}
\newcommand{\stCoalg}{{s_!t^*\textnormal{-coalg}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}

@@

\pietro{Qui o dopo.}
One of the key reasons for the success of neural networks is the notion of {\em weight sharing}. In its simplest form, one can consider, among maps $[M, X] \rightarrow [M, X]$, those that originate from maps $[M, X] \rightarrow X$. Indeed, given
\begin{equation*}
    {\cal F} \colon [M, X] \rightarrow X,
\end{equation*}
one can consider the map
\begin{align*}
    \widehat{\cal F} \colon [M, X]       & \rightarrow [M, X]              \\
    \left(\widehat{\cal F}\phi\right)(m) & = {\cal F}(a \mapsto \phi(ma)).
\end{align*}

@@

We aim to show that the simple notion of weight sharing defined above is sufficient to recover, among other examples, classical recurrent and convolutional neural networks. However, to do so, we will need to generalize not the weight sharing strategy, but rather the underlying category on which such strategy takes place.

@@

@@

Graph-like structures are a source of examples of nilpotent functors.

\begin{proposition}\label{prop:graph_architecture}
    Let $(E, V)$ be a directed acyclic graph. Let $\DCat$ be a category with finite products. The functor
    \begin{align*}
        F \colon \DCat^V & \rightarrow \DCat^V                  \\
        (F X)(v)         & = \prod_{u \, | \, (u, v)\in E} X(u)
    \end{align*}
    is nilpotent.
\end{proposition}

\begin{proof}
    Let $n$ be such that there are no paths of length $n$ in $(V, E)$. Then $F^n X$ is the terminal object.
\end{proof}

\Cref{prop:graph_architecture} has a continuous analogue in locally Cartesian closed categories, which will be our focus.

\begin{definition}\label{def:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. An internal quiver
    \begin{equation*}
        s, t\colon E \rightrightarrows V
    \end{equation*}
    {\em has bounded depth} if there exists $n \in \N$ such that the iterated pullback
    \begin{equation}\label{eq:nilpotent_left_adjoint}
        \underbrace{E \times_V E \times_V \dots \times_V E}_{n \text{ factors}}
    \end{equation}
    is initial in $\DCat/V$.
\end{definition}

\begin{proposition}\label{prop:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then \pietro{TODO: definisci * e ! (base change functors).}
    \begin{equation*}
        s_!t^*\colon \DCat/V \rightarrow \DCat/V
    \end{equation*}
    is a nilpotent endofunctor.
\end{proposition}

\begin{proof}
    The functor $s_!t^*$ is the right adjoint of the functor  $t_*s^*\colon \DCat/V \rightarrow \DCat/V$. By~\cref{eq:nilpotent_left_adjoint}, $(t_*s^*)^n$ maps all objects to the initial object, so its adjoint $(s_!t^*)^n$ maps all objects to the terminal object.
\end{proof}

@@

Let $\Cat$ be a Cartesian closed category with finite limits, and let $F\colon \Cat \rightarrow \Cat$ be a left exact (i.e., preserves finite limits) nilpotent endofunctor. We can consider $\FCoalg$, the category of coalgebras over $F$ \pietro{add ref}. Objects are simply maps
\begin{equation*}
    X \rightarrow F X,
\end{equation*}
where $X \in \Obj(\Cat)$. Morphisms are given by commutative diagrams
\begin{equation*}
    \begin{tikzcd}
        X \arrow[rightarrow]{d} \arrow[rightarrow]{r}{\phi}
        & Y \arrow[rightarrow]{d} \\
        FX \arrow[rightarrow]{r}{F \phi}
        & FY
    \end{tikzcd}
\end{equation*}
As $F$ is nilpotent and preserves finite products, it is straightforward to verify that the forgetful functor
\begin{equation*}
    U \colon \FCoalg \rightarrow \Cat
\end{equation*}
has a right adjoint $\Phi$, called the cofree functor, given by
\begin{equation*}
    \Phi(X) = X \times F X \times \dots \times F^{n-1} X \rightarrow F X \times \dots \times F^{n-1} X.
\end{equation*}

This adjunction is comonadic \pietro{explain! cite Beck's theorem!}, and $\FCoalg$ is the Eilenberg-Moore category $\Cat^T$, where $T = U\Phi$. As $T$ is left exact (product of left exact), $\FCoalg \simeq \Cat^T$ is Cartesian closed and it has finite limits, which are preserved by both $U$ and $\Phi$~\cite[Sect.~5.50]{wyler1991lecture}.

\begin{proposition}\label{prop:cofree_exponentiating}
    Let $F\colon \Cat \rightarrow \Cat$ be a left exact, nilpotent endofunctor. Then the exponential of cofree coalgebras is again cofree.
    More explicitly, let $\Phi\colon \Cat \rightarrow \FCoalg$ denote the cofree functor. Then, there exists a natural isomorphism
    \begin{equation*}
        [B, \Phi X] \simeq \Phi[UB, X].
    \end{equation*}
\end{proposition}

\begin{proof}
    By a straightforward computation:
    \begin{align*}
        \Hom_{\FCoalg}(A, [B, \Phi X])
         & \simeq \Hom_{\FCoalg}(A\times B, \Phi X)     \\
         & \simeq \Hom_{\Cat}(U(A \times B), X)  \\
         & \simeq \Hom_{\Cat}(UA \times UB, X)   \\
         & \simeq \Hom_{\Cat}(UA, [UB, X])       \\
         & \simeq \Hom_{\FCoalg}(A, \Phi[UB, X]).
    \end{align*}
    By Yoneda's lemma, this induces a natural isomorphism
    \begin{equation*}
        [B, \Phi X] \simeq \Phi[UB, X].
    \end{equation*}
\end{proof}

@@

% \begin{definition}\label{def:nilpotent_architecture}
%     A {\em nilpotent architecture} is a left exact, nilpotent endofunctor $F$ on a finitely complete, Cartesian closed category $\Cat$.
% \end{definition}

\begin{theorem}\label{thm:nilpotent_architecture}
    Let $F\colon \Cat \rightarrow \Cat$ be a left exact, nilpotent endofunctor. Let $\Phi(X)$ be a cofree coalgebra in $\FCoalg$. Then a map $F X \rightarrow X$ induces a machine $\Phi X \rightarrow \Phi X$.
\end{theorem}

\begin{proof}
    \pietro{TODO, da fare anche con parametric machines e $P \times F X \rightarrow X$, dove $P$ è una $F$ coalgebra con un'azione di $M$.}
\end{proof}

\begin{definition}\label{def:architecture}
    An {\em architecture} is a quiver of bounded depth $s,t \colon E \rightrightarrows V$ in a locally Cartesian closed category $\DCat$, together with a monoid object $M$ in $\stCoalg$.
\end{definition}

This is sufficient to give notions of composability, weight sharing, and locality.

\begin{theorem}\label{thm:weight_sharing}
    A nilpotent architecture $M, s, t$ induces a comonad on $\DCat/V$, given by
    \begin{equation*}
        X \mapsto U\Phi[UM, X].
    \end{equation*}
\end{theorem}

\begin{proof}
    \pietro{Free forget composed with $[M, \anon]$ in $\stCoalg$.}
\end{proof}

\pietro{Forse si possono riunificare le due nozioni!}

@@

Monoids can be considered as categories with one object. Let us consider the category corresponding to $\N$. Then the above construction recovers recurrent neural networks. Analogously, considering $\N \times \N$ we obtain convolutional neural networks \pietro{discuss shape of filter?}

>>>>

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Fun}{{\textnormal{Fun}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Set}{{\mathbf{Set}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\DCat}{{\mathbf{D}}}
\newcommand{\JCat}{{\mathcal{J}}}
\newcommand{\FCoalg}{{F\textnormal{-coalg}}}
\newcommand{\stCoalg}{{s_!t^*\textnormal{-coalg}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}

@@

One of the key reasons for the success of neural networks is the notion of {\em weight sharing}. In its simplest form, one can consider, among maps $[M, X] \rightarrow [M, X]$, those that originate from maps $[M, X] \rightarrow X$. Indeed, given
\begin{equation*}
    {\cal F} \colon [M, X] \rightarrow X,
\end{equation*}
one can consider the map
\begin{align*}
    \widehat{\cal F} \colon [M, X]       & \rightarrow [M, X]              \\
    \left(\widehat{\cal F}\phi\right)(m) & = {\cal F}(a \mapsto \phi(ma)).
\end{align*}

@@

We aim to show that the a similar strategy, based on a different comonad, is sufficient to recover, among other examples, classical recurrent and convolutional neural networks, incorporating at the same time equivariance and locality.

@@

\begin{definition}\label{def:cartesian_category}
    We say that a category is {\em Cartesian} if it has finite products. We say that a functor between Cartesian categories is {\em Cartesian} if it preserves finite products.
\end{definition}

\pietro{TODO definisci anche locally Cartesian closed!}

\begin{proposition}\label{prop:internal_category_comonad}
    Let $\DCat$ be a locally Cartesian closed category. Then, there is a well defined notion of category {\em internal} to $\DCat$~\cite{mac2013categories}.
    Let $\JCat = s, t \colon C_1 \rightrightarrows C_0$ be such an internal category. $\JCat$ induces a limit-preserving comonad
    \begin{equation*}
        t_!s^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0.
    \end{equation*}
    \pietro{TODO: definisci * e ! (base change functors).}
\end{proposition}

\begin{proof}
    The map $s_*t^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0$ is a monad~\cite[Thm.~V.8.2]{Mac_Lane_1994}. As $\DCat$ is locally Cartesian closed, $s^*$ has a right adjoint $s_!$, hence we can consider $t_!s^*$. As $T = t_!s^*$ is the right adjoint of the monad $s_*t^*$, it is a comonad. Being a right adjoint, $T$ preserves limits.
\end{proof}

@@

\begin{definition}\label{def:internal_quiver_bounded_depth}
    Let $\DCat$ be a locally Cartesian closed category. An internal quiver
    \begin{equation*}
        s, t\colon E \rightrightarrows V
    \end{equation*}
    {\em has bounded depth} if there exists $n \in \N$ such that the iterated pullback
    \begin{equation}\label{eq:nilpotent_left_adjoint}
        \underbrace{E \times_V E \times_V \dots \times_V E}_{n \text{ factors}}
    \end{equation}
    is initial in $\DCat/V$.
\end{definition}

\begin{proposition}\label{prop:internal_quiver_functor}
    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then
    \begin{equation*}
        s_!t^*\colon \DCat/V \rightarrow \DCat/V
    \end{equation*}
    is a limit-preserving nilpotent functor.
\end{proposition}

\begin{proof}
    The functor $s_!t^*$ is the right adjoint of the functor  $t_*s^*\colon \DCat/V \rightarrow \DCat/V$, hence it preserves limits and is therefore Cartesian. By~\cref{eq:nilpotent_left_adjoint}, $(t_*s^*)^n$ maps all objects to the initial object, so its adjoint $(s_!t^*)^n$ maps all objects to the terminal object.
\end{proof}

\section{Equivariant architectures}

Let us now assume that $\DCat$  is a locally Cartesian closed category with a terminal object (and thus all finite limits).

\begin{definition}\label{def:equivariance_category}
    Let $M$ be a monoid in $\DCat$ and $I \in \Obj$. The {\em equivariance category} $\JCat_{M, I}$ is given by the projection maps
    \begin{equation*}
        s, t\colon M \times I \times I \rightrightarrows I,
    \end{equation*}
    with composition and identity given by composition and identity in $M$. The {\em equivariance comonad} is the comonad induced by this category via~\cref{prop:internal_category_comonad}, that is, $t_!s^*$.
\end{definition}

We introduce the notion of {\em nilpotent constraint}, which we will need in order to limit equivariance and connectivity. To do so, we need some additional definitions.

\begin{definition}\label{def:nilpotent_architecture}
    Let $\JCat_{M, I}$ be an equivariance category. A {\em nilpotence constraint} on $\JCat_{M, I}$ is a subobject $S \subseteq M \times I \times I$ such that the internal quiver induced by
    \begin{equation*}
        S \hookrightarrow M \times I \times I \rightrightarrows I
    \end{equation*}
    has bounded depth, as in~\cref{def:internal_quiver_bounded_depth}.
\end{definition}
% Let $\Cat$ be a Cartesian category, and let $T$ be a comonad on $\Cat$. A {\em nilpotent constraint} on $T$ is a nilpotent Cartesian endofunctor $F \colon \Cat \rightarrow \Cat$ together with a natural transformation $\eta \colon T \rightarrow F$.

This is sufficient to give a notion of locality.

@@

Let $\Cat$ be a Cartesian closed category, and let $F\colon \Cat \rightarrow \Cat$ be a nilpotent architecture. We can consider $\FCoalg$, the category of coalgebras over $F$ \pietro{add ref}. Objects are simply maps
\begin{equation*}
    X \rightarrow F X,
\end{equation*}
where $X \in \Obj(\Cat)$. Morphisms are given by commutative diagrams
\begin{equation*}
    \begin{tikzcd}
        X \arrow[rightarrow]{d} \arrow[rightarrow]{r}{\phi}
        & Y \arrow[rightarrow]{d} \\
        FX \arrow[rightarrow]{r}{F \phi}
        & FY
    \end{tikzcd}
\end{equation*}
As $F$ is Cartesian, the category $\FCoalg$ is also Cartesian, with products given by
\begin{equation*}
    \prod_{i \in I} X_i \rightarrow \prod_{i \in I} F X_i \simeq F \prod_{i \in I} X_i.
\end{equation*}
As $F$ is nilpotent and preserves finite products, it is straightforward to verify that the forgetful functor
\begin{equation*}
    U \colon \FCoalg \rightarrow \Cat
\end{equation*}
has a right adjoint $\Phi$, called the cofree functor, given by
\begin{equation*}
    \Phi(X) = X \times F X \times \dots \times F^{n-1} X \rightarrow F X \times \dots \times F^{n-1} X.
\end{equation*}

\begin{remark}
    This adjunction is comonadic \pietro{explain! cite Beck's theorem!}, and $\FCoalg$ is the Eilenberg-Moore category $\Cat^T$, where $T = U\Phi$.
\end{remark}

@@

\begin{theorem}\label{thm:nilpotent_architecture}
    Let $F\colon \Cat \rightarrow \Cat$ be a nilpotent endofunctor. Let $\Phi X$ be a cofree coalgebra in $\FCoalg$. Let $P$ be a $F$ coalgebra. Then a map $P \times F X \rightarrow X$ induces a parametric machine $P \times \Phi X \rightarrow \Phi X$.
\end{theorem}

\begin{proof}
    \pietro{TODO per induzione.}
\end{proof}

\pietro{La sfida principale qui è trovare un buona architettura $F$ ma anche una buona coalgebra $P$. Probabilmente da menzionare da qualche parte che se la categoria di base è buona, $\FCoalg$ ha molte buone proprietà.}

@@

\pietro{discuss shape of filter?}

### 2020-10-08 backup/unify/main.tex
before volterra

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Fun}{{\textnormal{Fun}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Set}{{\mathbf{Set}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\DCat}{{\mathbf{D}}}
\newcommand{\JCat}{{\mathcal{J}}}
\newcommand{\FCoalg}{{F\textnormal{-coalg}}}
\newcommand{\stCoalg}{{s_!t^*\textnormal{-coalg}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}

@@

\section{Categorical preliminaries}

@@

\pietro{TODO definisci anche locally Cartesian closed!}

\begin{proposition}\label{prop:internal_category_comonad}
    Let $\DCat$ be a locally Cartesian closed category. Then, there is a well defined notion of category {\em internal} to $\DCat$~\cite{mac2013categories}.
    Let $\JCat = s, t \colon C_1 \rightrightarrows C_0$ be such an internal category. $\JCat$ induces a limit-preserving comonad
    \begin{equation*}
        t_!s^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0.
    \end{equation*}
    \pietro{TODO: definisci * e ! (base change functors).}
\end{proposition}

\begin{proof}
    The map $s_*t^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0$ is a monad~\cite[Thm.~V.8.2]{Mac_Lane_1994}. As $\DCat$ is locally Cartesian closed, $s^*$ has a right adjoint $s_!$, hence we can consider $t_!s^*$. As $T = t_!s^*$ is the right adjoint of the monad $s_*t^*$, it is a comonad. Being a right adjoint, $T$ preserves limits.
\end{proof}

@@

\begin{definition}\label{def:internal_quiver_bounded_depth}
    Let $\DCat$ be a locally Cartesian closed category. An internal quiver
    \begin{equation*}
        s, t\colon E \rightrightarrows V
    \end{equation*}
    {\em has bounded depth} if there exists $n \in \N$ such that the iterated pullback
    \begin{equation}\label{eq:nilpotent_left_adjoint}
        \underbrace{E \times_V E \times_V \dots \times_V E}_{n \text{ factors}}
    \end{equation}
    is initial in $\DCat/V$.
\end{definition}

\begin{proposition}\label{prop:internal_quiver_functor}
    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then
    \begin{equation*}
        s_!t^*\colon \DCat/V \rightarrow \DCat/V
    \end{equation*}
    is a limit-preserving nilpotent functor.
\end{proposition}

@@

\section{Equivariant architectures}

Let us now assume that $\DCat$  is a locally Cartesian closed category with a terminal object (and thus all finite limits).

\begin{definition}\label{def:equivariance_category}
    Let $M$ be a monoid in $\DCat$ and $I \in \Obj$. The {\em equivariance category} $\JCat_{M, I}$ is given by the projection maps
    \begin{equation*}
        s, t\colon M \times I \times I \rightrightarrows I,
    \end{equation*}
    with composition and identity given by composition and identity in $M$. The {\em equivariance comonad} is the comonad induced by this category via~\cref{prop:internal_category_comonad}, that is, $t_!s^*$.
\end{definition}

We introduce the notion of {\em nilpotent constraint}, which we will need in order to limit equivariance and connectivity. To do so, we need some additional definitions.

\begin{definition}\label{def:nilpotent_architecture}
    Let $\JCat_{M, I}$ be an equivariance category. A {\em nilpotence constraint} on $\JCat_{M, I}$ is a subobject $S \subseteq M \times I \times I$ such that the internal quiver induced by
    \begin{equation*}
        S \hookrightarrow M \times I \times I \rightrightarrows I
    \end{equation*}
    has bounded depth, as in~\cref{def:internal_quiver_bounded_depth}.
\end{definition}
% Let $\Cat$ be a Cartesian category, and let $T$ be a comonad on $\Cat$. A {\em nilpotent constraint} on $T$ is a nilpotent Cartesian endofunctor $F \colon \Cat \rightarrow \Cat$ together with a natural transformation $\eta \colon T \rightarrow F$.

This is sufficient to give a notion of locality.

>>>>

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Fun}{{\textnormal{Fun}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Set}{{\mathbf{Set}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\DCat}{{\mathbf{D}}}
\newcommand{\FCoalg}{{F\textnormal{-coalg}}}
\newcommand{\stCoalg}{{s_!t^*\textnormal{-coalg}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}

@@

\section{Architecture}

@@

@@

We begin by giving an abstract definition of {\em nilpotent architecture}.

\begin{definition}\label{def:nilpotent_architecture}
    Let $\Cat$ be a Cartesian category. A {\em nilpotent architecture} is a nilpotent Cartesian endofunctor $F \colon \Cat \rightarrow \Cat$.
\end{definition}

This is sufficient to give notions of composability, weight sharing, and locality.
Graph-like structures are a source of examples of nilpotent architecture.

\begin{proposition}\label{prop:graph_architecture}
    \pietro{Il multigrafo è più interessante, ma tanto è solo un esempio.}
    Let $(E, V)$ be a directed acyclic graph. Let $\DCat$ be a Cartesian category. The functor
    \begin{align*}
        F \colon \DCat^V & \rightarrow \DCat^V                  \\
        (F X)(v)         & = \prod_{u \, | \, (u, v)\in E} X(u)
    \end{align*}
    is a nilpotent architecture.
\end{proposition}

\begin{proof}
    It is straightforward to verify that $F$ is Cartesian.
    Let $n$ be such that there are no paths of length $n$ in $(V, E)$. Then $F^n X$ is the terminal object.
\end{proof}

\Cref{prop:graph_architecture} has a continuous analogue in locally Cartesian closed categories, which will be our focus.

\begin{definition}\label{def:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. An internal quiver
    \begin{equation*}
        s, t\colon E \rightrightarrows V
    \end{equation*}
    {\em has bounded depth} if there exists $n \in \N$ such that the iterated pullback
    \begin{equation}\label{eq:nilpotent_left_adjoint}
        \underbrace{E \times_V E \times_V \dots \times_V E}_{n \text{ factors}}
    \end{equation}
    is initial in $\DCat/V$.
\end{definition}

\begin{proposition}\label{prop:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then \pietro{TODO: definisci * e ! (base change functors).}
    \begin{equation*}
        s_!t^*\colon \DCat/V \rightarrow \DCat/V
    \end{equation*}
    is a nilpotent architecture.
\end{proposition}

@@
